---
title: "R for Data Science"
author: "Jongkyeong Kang"
date: "Kangwon National University"
output:
  beamer_presentation:
    latex_engine: pdflatex
    toc: no
    keep_tex: yes
    pandoc_args: --variable=mathspec
  pdf_document:
    toc: no
fontsize: 9pt
mathspec: yes
subtitle: Summary Statistics
theme: metropolis
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache =TRUE)
```

## Variable types

- When each entry in a vector comes from one of a small number of groups, we refer to the data as **categorical** data
  - sex (male or female)
  - US regions (Northeast, South, North Central, West)
  
- Some categorical data can be ordered (**ordinal data**)
   - spiciness(mild, medium, hot)

- Examples of numerical data:
   - population sizes, murder rates, and heights, etc.
- **Continuous variables** are those that can take any value, such as heights, if measured with enough precision.  
   - eg.  a pair of twins may be 68.12 and 68.11 inches, respectively.
-  Counts, such as number of gun murders per year, are **discrete** because they have to be round numbers
-  Discrete numeric data can be considered ordinal.

## Distributions

-  With categorical data, the distribution simply describes the proportion of each unique category.

```{r summaries-state-region-distribution}
prop.table(table(state.region))
table(state.region)/length(state.region)
```

- When the data is numerical, the task of constructing a summary based on the distribution is more challenging. 

## Case study: describing student heights

- Pretend that we have to describe the heights of our classmates to ET, an extraterrestrial that has never seen humans.

- We collect the heights of our students data and save it in the `heights`

```{r summaries-load-heights, warning=FALSE, message=FALSE}
library(tidyverse)
library(dslabs)
data(heights)
```

- One way to convey the heights to ET is to simply send him this list of 1050 heights.

-  But there are much more effective ways to convey this information, and understanding the concept of a distribution will help.

## Empirical cumulative distribution functions (경험적 누적 분포)

In general, when data is not categorical, reporting the frequency of each entry is not an effective summary since most entries are unique.
(수치데이터를 소수점까지 나타내는것은 적절하지 않다)
```{r}
height<-heights |> filter(sex=="Male") |> filter(height>=68 & height<69) |> dplyr::select(height) 
table(height)
```

-  empirical cumulative distribution function (eCDF) $F$
$$ F(a) = \mbox{Proportion of data points that are less than or equal to }a$$
ecdf함수를 통하여 그 값보다 작거나 같은 값으로 분포를 만든다.
## Empirical cumulative distribution functions

```{r summaries-ecdf, echo=FALSE, out.width="50%", out.height="50%", fig.align="center"}
ds_theme_set()  ## 없어도 코드에 문제는 없음
heights |> filter(sex=="Male") |> ggplot(aes(height)) + 
  stat_ecdf() +   ## stat_ecdf() ecdf를 그리는 함수
  ylab("F(a)") + xlab("a")
```

- $F(66)= 0.164$: 16\% of values are below 65
- $F(72)= 0.841$: 84\% of values are below 72
- $F(b)-F(a)$: the proportion of values between $a$ and $b$.

## Histograms

- Although the eCDF concept is widely discussed in statistics textbooks, the summary plot is actually not very popular in practice

   - At what value is the distribution centered?
   - Is the distribution symmetric?
   - What ranges contain 95% of the values?

- The simplest way to make a histogram is to divide the span of our data into non-overlapping bins of the same size. 

-  Then, for each bin, we count the number of values that fall in that interval. 

- The histogram plots these counts as bars with the base of the bar defined by the intervals.

## Histograms
```{r summaries-height-histogram, echo=FALSE, out.width="50%", out.height="50%", fig.align="center"}
heights |> 
  filter(sex=="Male") |> 
  ggplot(aes(height)) + 
  geom_histogram(binwidth = 1, color = "black")

height_M<-heights%>%filter(sex=="Male")
hist(height_M$height,breaks = seq(49.5,84.5,1))
```

 - The range of the data is from 50 to 84 with the majority (more than 95%) between 63 and 75 inches. 
 - The heights are close to symmetric around 69 inches.
 -  By adding up counts, ET could obtain a very good approximation of the proportion of the data in any interval. 
 
 -  Note that all values in each interval are treated the same when computing bin heights.
 
## Smoothed density

 Smooth density plots are similar to histograms, but the data is not divided into bins.
```{r summaries-example-of-smoothed-density, echo=FALSE, out.width="50%", out.height="50%", fig.align="center"}
heights |> 
  filter(sex=="Male") |> 
  ggplot(aes(height)) + 
  geom_density(alpha = .2, fill= "#00BFC4", color = 0)  +
  geom_line(stat='density')

A<-ggplot(height_M,aes(height))+
  geom_density(alpha=.4,fill="blue",color=0)
A+geom_line(stat="density")
```

- We no longer have sharp edges at the interval boundaries and many of the local peaks have been removed.
- The scale of the y-axis changed from counts to density.

## Smoothed density
### Hypothetical histogram with bins of size 1 with 2,000,000 values
```{r summaries--simulated-data-histogram-1, echo=FALSE, out.width="50%", out.height="50%", fig.align="center"}
set.seed(1988)
x <- data.frame(height = c(rnorm(1000000,69,3), rnorm(1000000,65,3)))
x |> ggplot(aes(height)) + geom_histogram(binwidth = 1, color = "black")

ggplot(data=x,aes(height))+ geom_histogram(binwidth = 1, color = "black")
```

## Smoothed density
### Hypothetical histogram with bins of size 1, 0.5, and 0.1 with 1,000,000 values
```{r summaries-simulated-data-histogram-2, fig.width=9, fig.height=3,  out.width = "100%",echo=FALSE, message=FALSE}
p1 <- x |> ggplot(aes(height)) + geom_histogram(binwidth = 1) + ggtitle("binwidth=1")
p2 <- x |> ggplot(aes(height)) + geom_histogram(binwidth = 0.5) + ggtitle("binwidth=0.5") 
p3 <- x |> ggplot(aes(height)) + geom_histogram(binwidth = 0.1) + ggtitle("binwidth=0.1")
library(gridExtra)
grid.arrange(p1, p2, p3, ncol = 3)
```
- The smooth density is basically the curve that goes through the top of the histogram bars when the bins are very, very small. 

## Smoothed density

```{r, simulated-density-1, echo=FALSE, out.width="50%", out.height="50%", fig.align="center"}
x |> ggplot(aes(height)) + 
  geom_histogram(aes(y=..density..), binwidth = 0.1) +
  geom_line(stat='density')

x |> ggplot(aes(height)) + 
  geom_histogram(aes(y=..density..), binwidth = 0.1) +
  geom_density(lwd=2,color=3,lintype=3)

```

- We don’t have millions of measurements. 
- Instead, we have 812 and we can’t make a histogram with very small bins.

- Make a histogram
   - using bin sizes appropriate for our data
   - computing frequencies rather than counts
   - drawing a smooth curve that goes through the tops of the histogram bars. 

## Smoothed density

```{r summaries-smooth-density-2, echo=FALSE, out.width = "100%"}
hist1 <- heights |> 
  filter(sex=="Male") |> 
  ggplot(aes(height)) +
  geom_histogram(aes(y=..density..), binwidth = 1, color="black") 
hist2 <- hist1 +
  geom_line(stat='density')
hist3 <- hist1 + 
  geom_point(data = ggplot_build(hist2)$data[[1]], aes(x,y), col = "blue")
hist4 <- ggplot() + geom_point(data = ggplot_build(hist2)$data[[1]], aes(x,y), col = "blue") + 
  xlab("height") + ylab("density")
hist5 <- hist4 + geom_line(data = ggplot_build(hist2)$data[[2]], aes(x,y))
hist6 <- heights |> 
  filter(sex=="Male") |> 
  ggplot(aes(height)) +
  geom_density(alpha = 0.2, fill="#00BFC4", col = 0) +
  geom_line(stat='density') +
  scale_y_continuous(limits = layer_scales(hist2)$y$range$range)
  
grid.arrange(hist1, hist3, hist4, hist5, hist2, hist6, nrow=2)
```


## Smoothed density
### Different smoothness
- We should select a degree of smoothness that we can defend as being representative of the underlying data.
  밀도의 대표값을 설정하므로서 선을 연결할수 있다.
```{r summaries-densities-different-smoothness, echo = FALSE, out.width = "100%", fig.width = 6, fig.height = 3}
p1 <- heights |> 
  filter(sex=="Male")|> ggplot(aes(height)) +
  geom_histogram(aes(y=..density..), binwidth = 1, alpha = 0.5) + 
  geom_line(stat='density', adjust = 0.5)
p2 <- heights |> 
  filter(sex=="Male") |> ggplot(aes(height)) +
  geom_histogram(aes(y=..density..), binwidth = 1, alpha = 0.5) + 
  geom_line(stat='density', adjust = 2)
grid.arrange(p1,p2, ncol=2)
```

## Smoothed density
### Proportion of values between 65 and 68
```{r summaries-area-under-curve, echo=FALSE, out.width="60%", out.height="60%", fig.align="center"}
d <- with(heights, density(height[sex=="Male"]))
tmp <- data.frame(height=d$x, density=d$y)
tmp |> ggplot(aes(height,density)) + geom_line() + 
  geom_area(aes(x=height,y=density), data = filter(tmp, between(height, 65, 68)), alpha=0.2, fill="#00BFC4")
```
- The proportion of this area is about 0.3, meaning that about 30% of male heights are between 65 and 68 inches.

## Normal distribution
- For any interval  $(a,b)$, the proportion of values in that interval can be computed using this formula:

$$\mbox{Pr}(a < x \leq b) = \int_a^b \frac{1}{\sqrt{2\pi}s} e^{-\frac{1}{2}\left( \frac{x-m}{s} \right)^2} \, dx$$

- It is completely defined by just two parameters: $m$ (mean) and $s$ (standard deviation). 

- The distribution is symmetric, centered at the average, and most values (about 95%) are within 2 SDs from the average.

## Normal distribution

For a list of numbers contained in a vector `x`, the average is defined as: 

```{r, eval=TRUE}
m <- sum(x) / length(x)
```

and the SD is defined as:
```{r}
s <- sqrt(sum((x-m)^2) / (length(x)-1))
```

Let's compute the values for the height for males which we will store in the object $x$:

```{r}
index <- heights$sex == "Male"
x <- heights$height[index]
```

The pre-built functions `mean` and `sd`  can be used here:
```{r}
m <- mean(x);s <- sd(x)
c(average = m, sd = s)
```

## Normal distribution
### normal distribution with mean 69.3 and SD =3.6(black line), student height smooth density (blue)
```{r summaries-data-and-normal-densities, echo=FALSE}
norm_dist <- data.frame(x = seq(-4, 4, len=50)*s + m) |> 
  mutate(density = dnorm(x, m, s))
heights |> filter(sex == "Male") |> ggplot(aes(height)) +
  geom_density(fill="#0099FF") +
  geom_line(aes(x, density),  data = norm_dist, lwd=1.5) 
```

## Standard units

- `z = (x-m)/s` with `m` and `s` the average and standard deviation of `x`, respectively.

- Why is this convenient?
  - The formula for the normal distribution is a function of `z` with `z = (x-m)/s`
  - The maximum of $e^{-z^2/2}$ is when $z=0$, which explains why the maximum of distribution occurs at the average
  - It also explains the symmetry
  - It reveals the relative position quickly. 

- In `R` we can obtain standard units using the function `scale`:
```{r}
z <- scale(x)
```

- How many men are within 2 SDs from the average? 95%신뢰구간
```{r}
mean(abs(z) < 2)
```
## Quantile-quantile plots

- A systematic way to assess how well the normal distribution fits the data is to check if the observed and predicted proportions match. 

- In general, this is the approach of the quantile-quantile plot (QQ-plot).

- $\Phi(x) = P(Z \le x)$, with $Z \sim N(0,1)$
  - $\Phi(-1.96) = 0.025$, $\Phi(1.96) = 0.975$. 

- In `R`, we can evaluate $\Phi$ using the `pnorm` function:

```{r}
pnorm(-1.96)
```

- The inverse function $\Phi^{-1}(x)$ gives us the _theoretical quantiles_ for the normal distribution. 
  - $\Phi^{-1}(0.975) = 1.96$. 
  
- In R, we can evaluate the inverse of $\Phi$ using the `qnorm` function.

```{r}
qnorm(0.975)
```

## Quantile-quantile plots

We can use `qnorm` to determine quantiles of a distribution with a specific average and standard deviation

```{r}
qnorm(0.975, mean = 5, sd = 2)
1.96*2+5
```

- If we have data in a vector $x$, we can define the quantile associated with any proportion $p$ as the $q$ for which the proportion of values below $q$ is $p$. 

- Using R code, we can define `q` as the value for which `mean(x <= q) = p`.

- Notice that not all $p$ have a $q$ for which the proportion is exactly $p$.

- To give a quick example, for the male heights data, we have that:
```{r}
mean(x <= 69.5)
```
So about 50% are shorter or equal to 69 inches. This implies that if $p=0.50$ then $q=69.5$.

## Quantile-quantile plots

 To construct a QQ-plot, we do the following:

1. Define a vector of $m$ proportions $p_1, p_2, \dots, p_m$.
2. Define a vector of quantiles $q_1, \dots, q_m$ for your data for the proportions $p_1, \dots, p_m$. We refer to these as the _sample quantiles_. 
3. Define a vector of theoretical quantiles for the proportions $p_1, \dots, p_m$ for a normal distribution with the same average and standard deviation as the data.
4. Plot the sample quantiles versus the theoretical quantiles.

## Quantile-quantile plots

- Start by defining the vector of proportions.
```{r}
p <- seq(0.05, 0.95, 0.05)
```

- To obtain the quantiles from the data, we can use the `quantile` function like this:
```{r}
sample_quantiles <- quantile(x, p)
head(sample_quantiles)
```

- To obtain the theoretical normal distribution quantiles with the corresponding average and SD, we use the `qnorm` function:
```{r}
theoretical_quantiles <- qnorm(p, mean = mean(x), sd = sd(x))
```

## Quantile-quantile plots

```{r summaries-qqplot-original}
qplot(theoretical_quantiles, sample_quantiles) + geom_abline()
```

```{r}
sample_quantiles2<-quantile(z,p)
theoretical_quantiles2<-qnorm(p)
plot(theoretical_quantiles,sample_quantiles)
abline(a=0,b=1)
```

## Quantile-quantile plots

- Notice that this code becomes much cleaner if we use standard units:
```{r summaries-qqplot-standardized,out.width="60%", out.height="60%", fig.align="center"}
sample_quantiles <- quantile(z, p)
theoretical_quantiles <- qnorm(p) 
qplot(theoretical_quantiles, sample_quantiles) + geom_abline()
```

## Quantile-quantile plots
- The above code is included to help describe QQ-plots. However, in practice it is easier to use **ggplot2** code:

```{r, out.width="60%", out.height="60%", fig.align="center"}
heights |> filter(sex == "Male") |>
  ggplot(aes(sample = scale(height))) + 
  geom_qq() +
  geom_abline()
```

## Quantile-quantile plots

- While for the illustration above we used 20 quantiles, the default from the `geom_qq` function is to use as many quantiles as data points.

- Note that although here we used qqplots to compare an observed distribution to the mathematically defined normal distribution, QQ-plots can be used to compare any two distributions.

## Percentiles 

- _Percentiles_ are special cases of  _quantiles_ that are commonly used.

- The percentiles are the quantiles you obtain when setting the $p$ at $0.01, 0.02, ..., 0.99$.

- We call, for example, the case of $p=0.25$ the 25th percentile, which gives us a number for which 25% of the data is below. 

- The most famous percentile is the 50th, also known as the _median_. 

- For the normal distribution the _median_ and average are the same, but this is generally not the case.

- Another special case that receives a name are the _quartiles_, which are obtained when setting $p=0.25,0.50$, and $0.75$. 

## Boxplots

- Suppose we want to summarize the murder rate(from US murders dataset) distribution. 

- Using the techniques we have learned, we can quickly see that the normal approximation does not apply here:

```{r summaries-hist-qqplot-non-normal-data, out.width = "100%",  fig.width = 6, fig.height = 3, echo=FALSE}
data(murders)
murders <- murders |> mutate(rate = total/population*100000)
library(gridExtra)
p1 <- murders |> ggplot(aes(x=rate)) + geom_histogram(binwidth = 0.5, color = "black") + ggtitle("Histogram")
p2 <- murders |> ggplot(aes(sample=rate)) + 
  geom_qq(dparams=summarize(murders, mean=mean(rate), sd=sd(rate))) +
  geom_abline() + ggtitle("QQ-plot")
grid.arrange(p1, p2, ncol = 2)
```

## Boxplots
- The boxplot provides a five-number summary composed of the range along with the quartiles (the 25th, 50th, and 75th percentiles). 

- The boxplot often ignore _outliers_ when computing the range and instead plot these as independent points. 

- Finally, he suggested we plot these numbers as a "box" with "whiskers" like this:

```{r summaries-first-boxplot, echo=FALSE, out.width="60%", out.height="60%", fig.align="center"}
murders |> ggplot(aes("",rate)) + geom_boxplot() +xlab("")
#  +coord_cartesian(xlim = c(0, 2))  차이가 없음
```

## Stratification {#stratification}

- In data analysis we often divide observations into groups based on the values of one or more variables associated with those observations.

- For example, we divide the height values into groups based on a sex variable: females and males. 

- We call this procedure _stratification_ and refer to the resulting groups as _strata_. 

- Stratification is common in data visualization because we are often interested in how the distribution of variables differs across different subgroups.

## Case study: describing student heights (continued)
### Boxplots of heigts for men and women

```{r summaries-female-male-boxplots, echo=FALSE}
heights |> ggplot(aes(x=sex, y=height, fill=sex)) +
  geom_boxplot()
```

## Case study: describing student heights (continued)
### Histogram and QQ-plot of heigts for women

```{r summaries-histogram-qqplot-female-heights, echo=FALSE, out.width="100%",  fig.width = 6, fig.height = 3}
p1 <- heights |> filter(sex == "Female") |>
  ggplot(aes(height)) +
  geom_density(fill="#F8766D") 
p2 <- heights |> filter(sex == "Female") |> 
  ggplot(aes(sample=scale(height))) +
  geom_qq() + geom_abline() + ylab("Standard Units")
grid.arrange(p1, p2, ncol=2)
```


## Case study: describing student heights (continued)

- The QQ-plot shows that the highest points tend to be taller than expected by the normal distribution. 

- We also see five points in the QQ-plot that suggest shorter than expected heights for a normal distribution. 

- Regarding the five smallest values, note that these values are:
```{r}
heights |> filter(sex == "Female") |> 
  top_n(5, desc(height)) |>
  pull(height)
```

- Because these are reported heights, a possibility is that the student meant to enter `5'1"`, `5'2"`, `5'3"` or `5'5"`. 
